---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# {devindocker}

<!-- badges: start -->
[![R build status](https://github.com/ThinkR-open/devindocker/workflows/R-CMD-check/badge.svg)](https://github.com/ThinkR-open/devindocker/actions)
<!-- badges: end -->

The goal of {devindocker}, as "Development In Docker", is to help launch your R project inside a Docker container with Rstudio server.  

- You can develop with the same architecture as your clients
- You can test your project / package on a specific architecture
- You can combine this with {renv} to keep packages versions installed inside your directory (and thus available next time you start the container)
- All changes in settings of your Rstudio Server will be kept for next start (last project opened, editor theme, ...)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("ThinkR-open/devindocker")
```

## Usage

Let us create a random directory with a file inside.
```{r}
# Temporary project
tempdir <- tempdir()
project_path <- file.path(tempdir, "myproject")
dir.create(project_path)
# Add a file inside
cat("# my R file", file = file.path(project_path, "my-file.R"))
```

### Work in a Docker environment 

Launch a Docker container with your directory inside. This should be a container with Rstudio server inside.  
_Note that packages you install will not be kept after you stop the container, but RStudio preferences will_
```{r}
# Which path to your working directory / project
project_path <- file.path(tempdir, "myproject")

# Which container (with Rstudio inside) ? ----
# https://hub.docker.com/r/rocker/verse
container <- "rocker/geospatial:4.0.1"

# Which port ? ----
# _Useful if multiple Rstudio Server to launch
port <- 8788

# Start Docker project ----
launch_proj_docker(
  project_path = project_path,
  container = container,
  port = port)
```

When you're done, do not forget to stop properly the Rstudio Server: Click on Top right button to quit or `q()` in the console.

Then, end the container.
```{r}
# Stop Docker properly
stop_proj_docker(project_path = project_path)
```

### Use {renv} inside Docker and keep installation of packages

**Note that you need to launch your project with {devindocker} from outside this project. Never ever open it again locally (out of a Docker container) if you want to avoid problems with bad and not compatible local {renv} setup. It is recommended to create a project dedicated to launch {devindocker} projects.**

Launch a Docker container with your directory inside. This should be a container with Rstudio server inside.  
_Note that packages you install will be kept after you stop the container, as well as RStudio preferences._

**Follow instructions in the `"renv_instructions.Rmd"` file that is created inside your project.**
```{r}
# Which path to your working directory / project
project_path <- file.path(tempdir, "myproject")

# Which container (with Rstudio inside) ? ----
# https://hub.docker.com/r/rocker/verse
container <- "rocker/geospatial:4.0.1"

# Which port ? ----
# _Useful if multiple Rstudio Server to launch
port <- 8788

# My renv cache directory on my local computer
# Used as persistent drive for all you Docker container with {devindocker}
renv_cache <- "~/renv_cache"

# Start Docker project ----
devindocker::launch_proj_docker(
  project_path = project_path,
  container = container,
  port = port,
  renv_cache = renv_cache,
  renv_inst = TRUE, # Add an Rmd with instructions inside your project
  update_docker = TRUE
)

```

When you're done, do not forget to stop properly the Rstudio Server: Click on Top right button to quit or `q()` in the console.

Then, end the container.
```{r}
# Stop Docker properly
stop_proj_docker(project_path = project_path)
```

