---
title: "Development in Docker with {renv}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ab-dev-in-docker-with-renv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(devindocker)
```

# Create new package

You can use {devindocker} for any directory, project or even package.
In the end, these are all folders on your computer.
```{r}
tempdir <- tempdir()
my_project <- normalizePath(file.path(tempdir, "myproject"), mustWork = FALSE)
usethis::create_package(my_project, open = FALSE)
```

### Use {renv} inside Docker and keep installation of packages

**Note that you need to launch your project with {devindocker} from outside this project. Never ever open it again locally (out of a Docker container) if you want to avoid problems with bad and not compatible local {renv} setup. It is recommended to create a project dedicated to launch {devindocker} projects.**

Launch a Docker container with your directory inside. 
This should be a container with Rstudio server.

Common and project {renv} libraries will be stored on your local computer (See details in the next section).

You will need to set up {renv} the first time you launch the Docker container.
**Follow instructions in the `"renv_instructions.Rmd"` file that is created inside your project.**

```{r, eval=FALSE}
# Which path to your working directory / project
project_path <- file.path(tempdir, "myproject")

# Which container (with Rstudio inside) ? ----
# https://hub.docker.com/r/rocker/verse
container <- "rocker/geospatial:4.0.1"

# Which port ? ----
# _Useful if multiple Rstudio Server to launch
port <- 8788

# My renv cache directory on my local computer
# Used as persistent drive for all you Docker container with {devindocker}
renv_cache <- "~/renv_cache"

# Start Docker project ----
devindocker::launch_proj_docker(
  project_path = project_path,
  container = container,
  port = port,
  renv_cache = renv_cache,
  renv_inst = TRUE, # Add an Rmd with instructions inside your project
  update_docker = TRUE
)

# Follow instructions in "renv_instructions.Rmd" to setup {renv}
```

When you're done, do not forget to stop properly the Rstudio Server: Click on Top right button to quit or `q()` in the console.

Then, stop the container.
```{r, eval=FALSE}
# Stop Docker properly
stop_proj_docker(project_path = project_path)
```

### {renv} and its different storage directories

{renv} has two levels of storing downloaded packages. 
First one is for the project, second one is for your computer:

- `".renv"` directory for the project is by default inside the project directory itself.
This has some consequences when your project is a package during `check()` because it is copied in the temporary directory of the `check()`.
A big project library may take some time and some place in your Docker container.
  + Using {devindocker}, these project specific packages are stored locally in your `project_path` directory by default, which is the default {renv} behaviour. 
Indeed, if you use global cache (with `renv_cache`), only symlinks are stored in these directory.  
You can change it with parameters `renv_out = TRUE` to save these in a folder called `".renv"`, in the parent directory of your project `renv_out_dir`. 
All project stored in the same parent directory will share this local `".renv"` library, using a specific folder inside.
Inside the Docker container, the `".renv"` directory will be stored at the root of your RStudio server, in the Docker home directory. 
Show hidden files to se it.
- The directory for your computer will stored all versions of all packages downloaded, whether you continue to use them in some projects or not.
This directory is classified by R versions and packages.
In projects using {renv}, when you install a new package, {renv} will use this common directory to first check if you did not already downloaded the package.
Then, it will link the specific desired version of the computer library to the project library wih a symlink.
Inside Docker there is also a common directory.
Using {devindocker}, this common library will be linked to your local computer library that you can specify using `renv_cache` paramter.

Hence, 

- Packages you install for your project in the Docker will be kept after you stop the container, using a mounted volume. 
- Global package repository common to all your {renv} project will be shared with your computer, using a mounted volume
- RStudio preferences will be kept inside your project thanks to internal links with the Docker and a mounted volume.

## Activate cache shared on host
- Store this code in your *project* `.Rprofile` file
  `Sys.setenv(RENV_PATHS_CACHE = "/opt/local/renv/cache")`
- If cache *not* shared with host with option `renv_cache`, then de-activate cache otherwise packages will disappear at next Docker restart
`renv::settings$use.cache(FALSE)`
