---
title: "Include the Docker container in a network"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ac-docker-network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script type="text/javascript">
$(document).ready(function () {
  window.initializeCodeFolding("show" === "show");
});
</script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(devindocker)
```

To add a Docker container in a network where there is a database container for instance, you can use parameter `network_name` in function `launch_proj_docker()`.

## Connexion with a mysql container

Here is an example on how to start a Docker container with a MySQL database inside a network named `r-db`.  
*Note that your RStudio Container will then need additional system dependencies to be able to connect to the database. See examples in https://github.com/ColinFay/r-db .*

```{r}
network_name <- "r-db"

mysql_docker <- "mysql:8.0.16"
path_to_persistant_db <- "~/my_mysql_db"

# Databases container ----
## Pull image
system(paste("docker pull", mysql_docker))
## Create persistant drive for db
db_path <- normalizePath(path_to_persistant_db, mustWork = FALSE)
if (!dir.exists(db_path)) {
  dir.create(db_path)
}
## Create docker network
system(paste("docker network create", network_name))
## Start container
future::future({
  system(
    paste0(
      'docker run --net ', network_name,
      ' --name mysql ',
      ' -v ', db_path, ':/var/lib/mysql',
      ' -e MYSQL_ROOT_PASSWORD=coucou -d ', mysql_docker,
      ' --secure-file-priv=""',
      ' --default-authentication-plugin=mysql_native_password',
      ' && sleep 10',
      ' && docker exec mysql mysql -uroot -pcoucou -e "create database mydb" &')
  )
})

Sys.sleep(9)
```

Then, you can start the RStudio container (with database system dependencies) using {devindocker}

```{r}
tempdir <- tempdir()
my_project <- normalizePath(file.path(tempdir, "myproject"), mustWork = FALSE)
usethis::create_package(my_project, open = FALSE)
```

## Launch Docker
```{r, eval=FALSE}
# project_path <- "" # default
project_path <- my_project

# Which container with RStudio server? ----
container <- "thinkr/rstudio3_6_1_geo"

# Which port ? ----
# _Useful if multiple Rstudio Server to launch
port <- 8788

# Start Docker project
launch_proj_docker(project_path = project_path,
                   network_name = network_name,
                   container = container,
                   port = port)
```

## Connect to database inside the container

Inside the container, you can connect to your database using {DBI} or {dbplyr}.

```r
library(DBI)

con <- DBI::dbConnect(
  RMariaDB::MariaDB(), 
  user = "root",
  password = "coucou",
  host = "mysql", 
  db = "mydb"
)


dbListTables(con)
dbWriteTable(con, "mtcars", mtcars)
dbListTables(con)

res <- dbSendQuery(con, "SELECT * FROM mtcars WHERE cyl = 4")
dbFetch(res)
```


## Stop your db container

```{r}
system("docker kill mysql")
system("docker rm mysql")
```

## Stop RStudio Docker container and the network

```{r}
# Stop Docker and network properly
stop_proj_docker(project_path = project_path, sleep = 5,
                 network_name = network_name, stop_network = TRUE)
```

